% 论文基本信息
\ctitle{基于驱动程序替换的物联网设备固件仿真技术研究}
\etitle{Research on IoT Device Firmware Emulation Technology Based on Driver Replacement}

% 中文摘要
\cabstract{
随着物联网技术的快速发展，物联网设备数量呈现爆炸式增长，其固件安全问题日益突出。固件仿真技术是进行固件安全测试和漏洞挖掘的重要手段，然而现有的固件仿真方法存在外设建模成本高、硬件依赖强、自动化程度不足等问题。传统的基于硬件建模的方法需要为每种外设手动建立仿真模型，工作量大且难以覆盖多样化的硬件生态；基于驱动层替换的方法虽然避免了底层硬件建模，但驱动函数的定位与替换仍高度依赖人工干预，难以实现规模化应用。

针对上述问题，本文提出一种结合大语言模型的自动化固件仿真方法。该方法的核心思想是通过静态分析自动识别固件中的硬件依赖代码，利用大语言模型的代码理解与生成能力自动生成可替代的驱动实现，构建动态测试反馈闭环以迭代优化仿真质量，从而在无需手动建立外设模型的情况下实现固件在仿真环境中的稳定运行。

本文的主要工作包括三个方面。首先，设计并实现了面向嵌入式固件的硬件依赖代码自动识别模块，该模块通过静态分析技术，从驱动程序源码中自动提取硬件相关的函数调用点、外设寄存器访问模式和驱动函数边界，构建可查询的驱动知识库。其次，构建了基于大语言模型的驱动函数分析与替换模块，该模块利用大语言模型理解驱动函数的语义，通过结构化提示词模板和函数分类策略，自动生成可在仿真环境中运行的替代代码，并通过编译修复循环保障代码的可编译性。最后，引入了动态测试反馈闭环机制，当固件在仿真环境中运行出现异常时，系统能够自动分析异常原因并反馈给驱动替换模块，通过"生成—测试—反馈—修正"的闭环流程持续优化仿真质量。

实验结果表明，本文方法可支持多种嵌入式平台（STM32、NXP、Atmel）和操作系统（FreeRTOS、RT-Thread、Zephyr、裸机）的固件仿真，在给定实验设置下降低了固件仿真的技术门槛和人工成本，为漏洞挖掘奠定了基础。
}

% 中文关键词
\ckeywords{物联网安全；固件仿真；驱动程序替换；大语言模型；静态分析}

% 英文摘要
\eabstract{
With the rapid development of Internet of Things (IoT) technology, the number of IoT devices has grown explosively, and firmware security issues have become increasingly prominent. Firmware emulation is an important approach for firmware security testing and vulnerability discovery. However, existing firmware emulation methods face challenges such as high peripheral modeling costs, strong hardware dependencies, and insufficient automation. Traditional hardware modeling-based methods require manually building simulation models for each peripheral, which is labor-intensive and difficult to cover the diverse hardware ecosystem. Although driver layer replacement-based methods avoid low-level hardware modeling, the identification and replacement of driver functions still heavily rely on manual intervention, making it difficult to achieve large-scale application.

To address these issues, this paper proposes an automated firmware emulation method combined with large language models (LLMs). The core idea is to automatically identify hardware-dependent code in firmware through static analysis, use the code understanding and generation capabilities of LLMs to automatically generate alternative driver implementations, and build a dynamic testing feedback loop to iteratively optimize emulation quality, thereby enabling firmware to run stably in emulation environments without manually building peripheral models.

The main contributions of this paper include three aspects. First, a hardware dependency code automatic identification module for embedded firmware is designed and implemented, which automatically extracts hardware-related function call points, peripheral register access patterns, and driver function boundaries from firmware through MMIO pattern recognition, data flow tracking, and recursive induction techniques, constructing a queryable driver knowledge base. Second, a driver function analysis and replacement module based on LLMs is built, which utilizes LLMs to understand driver function semantics, automatically generates alternative code that can run in emulation environments through structured prompt templates and function classification strategies, and ensures code compilability through a compilation repair loop. Finally, a dynamic testing feedback loop mechanism is introduced. When firmware runs abnormally in the emulation environment, the system can automatically analyze the cause of the abnormality and feed it back to the driver replacement module, continuously optimizing emulation quality through a "generate-test-feedback-revise" closed-loop process.

Experimental results show that the proposed method supports firmware emulation across multiple embedded platforms (STM32, NXP, Atmel) and operating systems (FreeRTOS, RT-Thread, Zephyr, bare metal), and under the given experimental setup reduces the technical barrier and manual cost of firmware emulation, laying the foundation for vulnerability discovery.
}

% 英文关键词
\ekeywords{IoT Security; Firmware Emulation; Driver Replacement; Large Language Model; Static Analysis}
